---
title: "River Remedy: Comprehensive Contamination Analysis"
subtitle: "Pilcomayo River Basin Heavy Metal Assessment (2006-2024)"
author: "River Remedy Team"
date: "`r Sys.Date()`"
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    smooth-scroll: true
    fig-width: 12
    fig-height: 8
    embed-resources: true
    link-external-newwindow: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
---

```{r setup}
#| include: false

# Load all required libraries
library(tidyverse)
library(readxl)
library(janitor)
library(ggplot2)
library(scales)
library(viridis)
library(patchwork)
library(DT)
library(knitr)
library(leaflet)
library(plotly)

# Source analysis scripts
source("src/scripts/load_standards.R")

# Load standards
standards <- load_regulatory_standards()

# Define color palettes
risk_colors <- c(
  "Safe" = "#2E8B57", "Moderate" = "#FFD700", "High" = "#FF6347", 
  "Critical" = "#DC143C", "Extreme" = "#8B0000"
)

study_colors <- c(
  "2006" = "#4A90E2", "2011" = "#F5A623", "2024" = "#7ED321"
)

# Enhanced theme
theme_river_remedy <- function() {
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, size = 12, color = "gray60"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
}

# Risk calculation functions
calc_risk <- function(measured, limit) {
  ifelse(is.na(measured) | is.na(limit) | limit <= 0, "Unknown",
    ifelse(measured / limit <= 1, "Safe",
      ifelse(measured / limit <= 2, "Moderate",
        ifelse(measured / limit <= 10, "High",
          ifelse(measured / limit <= 100, "Critical", "Extreme")))))
}

# Data preparation function
prepare_data <- function(file_path, study_period, matrix_type) {
  if (!file.exists(file_path)) {
    return(NULL)
  }
  
  tryCatch({
    raw_data <- read_csv(file_path, show_col_types = FALSE) %>%
      clean_names()
    
    # Auto-detect metal columns
    metal_cols <- names(raw_data)[str_detect(names(raw_data), 
                                           "pb|as|cd|hg|cu|zn|fe|mn|\\(mg")]
    
    if (length(metal_cols) == 0) {
      return(NULL)
    }
    
    # Convert to long format
    processed <- raw_data %>%
      select(location = 1, all_of(metal_cols)) %>%
      pivot_longer(
        cols = all_of(metal_cols),
        names_to = "parameter", 
        values_to = "measured_value"
      ) %>%
      mutate(
        metal = case_when(
          str_detect(parameter, "pb|plomo") ~ "Lead",
          str_detect(parameter, "as|arsenico") ~ "Arsenic",
          str_detect(parameter, "cd|cadmio") ~ "Cadmium",
          str_detect(parameter, "hg|mercurio") ~ "Mercury", 
          str_detect(parameter, "cu|cobre") ~ "Copper",
          str_detect(parameter, "zn|zinc") ~ "Zinc",
          str_detect(parameter, "fe|hierro") ~ "Iron",
          str_detect(parameter, "mn|manganeso") ~ "Manganese",
          TRUE ~ "Unknown"
        ),
        measured_value = as.numeric(measured_value),
        study_period = study_period,
        matrix_type = matrix_type
      ) %>%
      filter(!is.na(measured_value), measured_value > 0, metal != "Unknown") %>%
      select(location, metal, measured_value, study_period, matrix_type)
    
    return(processed)
    
  }, error = function(e) {
    return(NULL)
  })
}

# Load all data
data_2006_list <- list(
  water = prepare_data("data/raw/ITA_water_2006.csv", "2006", "Water"),
  soil = prepare_data("data/raw/ITA_soil_2006.csv", "2006", "Soil"),
  sediment = prepare_data("data/raw/ITA_sed_2006.csv", "2006", "Sediment"),
  fish = prepare_data("data/raw/ITA_fish_2006.csv", "2006", "Fish"),
  human = prepare_data("data/raw/ITA_human_2006.csv", "2006", "Human Blood"),
  animal = prepare_data("data/raw/ITA_animal_2006.csv", "2006", "Animal Blood")
)

# Combine all data
all_data_list <- data_2006_list[!map_lgl(data_2006_list, is.null)]
all_data <- bind_rows(all_data_list)

# Perform risk assessment
risk_assessment <- all_data %>%
  rowwise() %>%
  mutate(
    who_standard = get_standard_value(metal, "who_water", standards),
    exceedance_ratio = measured_value / who_standard,
    risk_level = calc_risk(measured_value, who_standard)
  ) %>%
  ungroup()
```

# Executive Summary

::: callout-important
## 🚨 Critical Findings

This comprehensive analysis of the Pilcomayo River Basin reveals **catastrophic heavy metal contamination** with:

-   **90.4% of samples exceeding WHO standards**
-   **Lead contamination up to 231,150x WHO limits**
-   **Multiple extreme contamination hotspots identified**
-   **Systemic contamination across all environmental matrices**
:::

```{r summary-stats}
#| include: true

# Calculate key summary statistics
total_samples <- nrow(risk_assessment)
locations <- length(unique(risk_assessment$location))
matrices <- length(unique(risk_assessment$matrix_type))
metals <- length(unique(risk_assessment$metal))

risk_summary <- risk_assessment %>%
  count(risk_level) %>%
  mutate(percentage = round(n / sum(n) * 100, 1))

exceeding_standards <- sum(risk_summary$percentage[risk_summary$risk_level != "Safe"])

# Create summary cards
summary_cards <- tibble(
  Metric = c("Total Samples", "Locations Studied", "Sample Matrices", "Heavy Metals", "Exceeding WHO Standards"),
  Value = c(
    format(total_samples, big.mark = ","),
    as.character(locations),
    as.character(matrices), 
    as.character(metals),
    paste0(exceeding_standards, "%")
  ),
  Icon = c("📊", "🗺️", "🧪", "⚗️", "🚨")
)

kable(summary_cards, format = "html", escape = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

# Study Overview

This report analyzes heavy metal contamination across the Pilcomayo River Basin using data from the **2006 ITA Baseline Study**, which sampled `r total_samples` observations across `r locations` locations in `r matrices` different environmental matrices.

## Sample Distribution

```{r sample-distribution}
sample_dist <- all_data %>%
  count(matrix_type, metal) %>%
  ggplot(aes(x = matrix_type, y = n, fill = metal)) +
  geom_col(position = "stack", alpha = 0.8) +
  scale_fill_viridis_d(name = "Heavy Metal") +
  labs(
    title = "Sample Distribution by Matrix and Metal",
    subtitle = "Total samples across all environmental matrices",
    x = "Sample Matrix",
    y = "Number of Samples"
  ) +
  theme_river_remedy() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(sample_dist)
```

# Contamination Risk Assessment

## Overall Risk Distribution

```{r risk-overview}
risk_dist <- risk_assessment %>%
  count(risk_level) %>%
  mutate(
    percentage = round(n / sum(n) * 100, 1),
    risk_level = factor(risk_level, levels = names(risk_colors))
  ) %>%
  ggplot(aes(x = risk_level, y = percentage, fill = risk_level)) +
  geom_col(alpha = 0.8, width = 0.7) +
  geom_text(aes(label = paste0(percentage, "%\n(", n, ")")), 
            vjust = -0.5, fontface = "bold", size = 4) +
  scale_fill_manual(values = risk_colors, name = "Risk Level") +
  labs(
    title = "Risk Level Distribution - All Samples",
    subtitle = "Percentage and count of samples in each risk category",
    x = "Risk Level",
    y = "Percentage of Samples (%)",
    caption = "Based on WHO drinking water standards"
  ) +
  theme_river_remedy() +
  theme(legend.position = "none") +
  ylim(0, max(risk_summary$percentage) * 1.15)

print(risk_dist)
```

::: callout-warning
## Key Risk Statistics

```{r risk-table}
risk_summary %>%
  mutate(
    Risk_Level = risk_level,
    Samples = n,
    Percentage = paste0(percentage, "%")
  ) %>%
  select(Risk_Level, Samples, Percentage) %>%
  kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::row_spec(which(risk_summary$risk_level == "Critical"), background = "#DC143C", color = "white") %>%
  kableExtra::row_spec(which(risk_summary$risk_level == "Extreme"), background = "#8B0000", color = "white")
```
:::

## Risk by Sample Matrix

```{r risk-by-matrix}
matrix_risk <- risk_assessment %>%
  count(matrix_type, risk_level) %>%
  ggplot(aes(x = matrix_type, y = n, fill = risk_level)) +
  geom_col(position = "stack", alpha = 0.8) +
  scale_fill_manual(values = risk_colors, name = "Risk Level") +
  labs(
    title = "Risk Distribution by Sample Matrix",
    subtitle = "Number of observations in each risk category by sample type",
    x = "Sample Matrix",
    y = "Number of Observations"
  ) +
  theme_river_remedy() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(matrix_risk)
```

# Lead Contamination Analysis

Lead shows the most severe contamination across all matrices. The following analysis focuses on lead as the priority contaminant.

```{r lead-analysis}
lead_data <- risk_assessment %>%
  filter(metal == "Lead") %>%
  arrange(desc(exceedance_ratio))

# Top 20 lead contamination cases
top_lead <- lead_data %>%
  head(20) %>%
  ggplot(aes(x = reorder(paste(location, matrix_type, sep = " - "), exceedance_ratio),
             y = exceedance_ratio,
             fill = risk_level)) +
  geom_col(alpha = 0.8) +
  geom_hline(yintercept = 1, color = "red", linetype = "dashed", linewidth = 1.2) +
  scale_fill_manual(values = risk_colors, name = "Risk Level") +
  scale_y_continuous(
    trans = "log10", 
    labels = function(x) paste0(scales::comma(x, accuracy = 1), "x"),
    expand = expansion(mult = c(0, 0.1))
  ) +
  coord_flip() +
  labs(
    title = "Top 20 Lead Contamination Cases",
    subtitle = "Exceedance ratios vs WHO drinking water standard (0.01 mg/L)",
    x = "Location - Sample Type",
    y = "Exceedance Ratio (log scale)",
    caption = "Red line = WHO limit | Higher values indicate more severe contamination"
  ) +
  theme_river_remedy()

print(top_lead)
```

## Lead Contamination by Location

```{r lead-geographic}
lead_geographic <- lead_data %>%
  group_by(location) %>%
  summarise(
    n_samples = n(),
    avg_exceedance = mean(exceedance_ratio, na.rm = TRUE),
    max_exceedance = max(exceedance_ratio, na.rm = TRUE),
    matrices_affected = paste(unique(matrix_type), collapse = ", "),
    .groups = "drop"
  ) %>%
  arrange(desc(max_exceedance)) %>%
  head(15)

lead_location_plot <- lead_geographic %>%
  ggplot(aes(x = reorder(location, max_exceedance), y = max_exceedance, fill = log10(max_exceedance))) +
  geom_col(alpha = 0.8) +
  scale_fill_viridis_c(name = "Log₁₀\nExceedance", option = "plasma") +
  scale_y_continuous(
    trans = "log10",
    labels = function(x) paste0(scales::comma(x, accuracy = 1), "x")
  ) +
  coord_flip() +
  labs(
    title = "Lead Contamination by Geographic Location",
    subtitle = "Maximum exceedance ratios vs WHO standards",
    x = "Location",
    y = "Maximum Exceedance Ratio (log scale)"
  ) +
  theme_river_remedy()

print(lead_location_plot)
```

## Critical Contamination Sites

```{r critical-sites-table}
critical_sites <- risk_assessment %>%
  filter(risk_level %in% c("Critical", "Extreme")) %>%
  arrange(desc(exceedance_ratio)) %>%
  head(20) %>%
  mutate(
    exceedance_ratio = round(exceedance_ratio, 0),
    measured_value = round(measured_value, 3)
  ) %>%
  select(location, metal, matrix_type, measured_value, exceedance_ratio, risk_level)

DT::datatable(
  critical_sites,
  options = list(
    pageLength = 20,
    scrollX = TRUE,
    dom = 'Bfrtip'
  ),
  caption = "Top 20 Critical Contamination Cases",
  colnames = c("Location", "Metal", "Sample Type", "Concentration (mg/L)", "WHO Exceedance", "Risk Level")
) %>%
  formatStyle(
    "risk_level",
    backgroundColor = styleEqual(
      c("Critical", "Extreme"),
      c("#DC143C", "#8B0000")
    ),
    color = "white",
    fontWeight = "bold"
  ) %>%
  formatStyle(
    "exceedance_ratio",
    color = styleInterval(c(100, 1000, 10000), c("black", "orange", "red", "darkred")),
    fontWeight = "bold"
  )
```

# Multi-Metal Analysis

## Metal Priority Ranking

```{r metal-priority}
metal_priority <- risk_assessment %>%
  group_by(metal) %>%
  summarise(
    n_locations = n_distinct(location),
    n_samples = n(),
    avg_exceedance = mean(exceedance_ratio, na.rm = TRUE),
    max_exceedance = max(exceedance_ratio, na.rm = TRUE),
    critical_extreme_count = sum(risk_level %in% c("Critical", "Extreme")),
    priority_score = log10(avg_exceedance + 1) * n_locations,
    .groups = "drop"
  ) %>%
  arrange(desc(priority_score)) %>%
  mutate(
    avg_risk_category = case_when(
      avg_exceedance <= 1 ~ "Safe",
      avg_exceedance <= 2 ~ "Moderate",
      avg_exceedance <= 10 ~ "High",
      avg_exceedance <= 100 ~ "Critical",
      TRUE ~ "Extreme"
    )
  )

priority_plot <- metal_priority %>%
  ggplot(aes(x = reorder(metal, priority_score), y = priority_score, fill = avg_risk_category)) +
  geom_col(alpha = 0.8) +
  scale_fill_manual(values = risk_colors, name = "Average Risk") +
  coord_flip() +
  labs(
    title = "Metal Contamination Priority Ranking",
    subtitle = "Priority score = log(avg exceedance + 1) × number of locations",
    x = "Heavy Metal",
    y = "Priority Score",
    caption = "Higher scores indicate greater priority for intervention"
  ) +
  theme_river_remedy()

print(priority_plot)
```

## Metal Priority Table

```{r metal-table}
priority_table <- metal_priority %>%
  mutate(
    avg_exceedance = round(avg_exceedance, 1),
    max_exceedance = round(max_exceedance, 0),
    priority_score = round(priority_score, 2)
  ) %>%
  select(metal, n_samples, avg_exceedance, max_exceedance, critical_extreme_count, avg_risk_category, priority_score)

DT::datatable(
  priority_table,
  options = list(pageLength = 10),
  caption = "Metal Contamination Priority Analysis",
  colnames = c("Metal", "Samples", "Avg Exceedance", "Max Exceedance", "Critical/Extreme Cases", "Avg Risk", "Priority Score")
) %>%
  formatStyle(
    "avg_risk_category",
    backgroundColor = styleEqual(
      names(risk_colors),
      risk_colors
    ),
    color = "white",
    fontWeight = "bold"
  )
```

# Geographic Hotspot Analysis

```{r geographic-hotspots}
geographic_summary <- risk_assessment %>%
  group_by(location) %>%
  summarise(
    n_samples = n(),
    metals_tested = n_distinct(metal),
    matrices_tested = n_distinct(matrix_type),
    avg_exceedance = mean(exceedance_ratio, na.rm = TRUE),
    max_exceedance = max(exceedance_ratio, na.rm = TRUE),
    critical_extreme_count = sum(risk_level %in% c("Critical", "Extreme")),
    contamination_severity = case_when(
      max_exceedance >= 1000 ~ "Extreme Hotspot",
      max_exceedance >= 100 ~ "Critical Hotspot", 
      max_exceedance >= 10 ~ "High Contamination",
      max_exceedance >= 1 ~ "Moderate Contamination",
      TRUE ~ "Low Contamination"
    ),
    .groups = "drop"
  ) %>%
  arrange(desc(max_exceedance))

hotspot_plot <- geographic_summary %>%
  head(15) %>%
  ggplot(aes(x = reorder(location, max_exceedance), y = max_exceedance, fill = contamination_severity)) +
  geom_col(alpha = 0.8) +
  scale_fill_manual(
    values = c(
      "Extreme Hotspot" = "#8B0000",
      "Critical Hotspot" = "#DC143C", 
      "High Contamination" = "#FF6347",
      "Moderate Contamination" = "#FFD700",
      "Low Contamination" = "#2E8B57"
    ),
    name = "Contamination\nSeverity"
  ) +
  scale_y_continuous(
    trans = "log10",
    labels = function(x) paste0(scales::comma(x, accuracy = 1), "x")
  ) +
  coord_flip() +
  labs(
    title = "Geographic Contamination Hotspots",
    subtitle = "Maximum exceedance ratios by location (top 15)",
    x = "Location",
    y = "Maximum Exceedance Ratio (log scale)"
  ) +
  theme_river_remedy()

print(hotspot_plot)
```

# Standards Comparison

## Regulatory Framework

This analysis compares contamination levels against multiple regulatory standards:

```{r standards-reference}
standards_ref <- standards$reference_values %>%
  select(metal, who_water_mg_l, codex_food_mg_kg, epa_soil_mg_kg, cdc_blood_ug_dl) %>%
  rename(
    "Metal" = metal,
    "WHO Water (mg/L)" = who_water_mg_l,
    "Codex Food (mg/kg)" = codex_food_mg_kg, 
    "EPA Soil (mg/kg)" = epa_soil_mg_kg,
    "CDC Blood (µg/dL)" = cdc_blood_ug_dl
  )

DT::datatable(
  standards_ref,
  options = list(pageLength = 15, scrollX = TRUE),
  caption = "Regulatory Standards Reference Table"
) %>%
  formatRound(columns = 2:5, digits = 4)
```

# Environmental Impact Assessment

## Multi-Matrix Contamination

```{r multimatrix-heatmap}
# Create contamination heatmap
heatmap_data <- risk_assessment %>%
  group_by(location, matrix_type) %>%
  summarise(
    avg_exceedance = mean(exceedance_ratio, na.rm = TRUE),
    max_exceedance = max(exceedance_ratio, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    risk_category = case_when(
      max_exceedance <= 1 ~ "Safe",
      max_exceedance <= 2 ~ "Moderate",
      max_exceedance <= 10 ~ "High",
      max_exceedance <= 100 ~ "Critical", 
      TRUE ~ "Extreme"
    )
  )

# Select top locations for readability
top_locations <- geographic_summary$location[1:10]

heatmap_plot <- heatmap_data %>%
  filter(location %in% top_locations) %>%
  ggplot(aes(x = matrix_type, y = reorder(location, max_exceedance), fill = risk_category)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(
    aes(label = ifelse(max_exceedance > 10, 
                      paste0(round(max_exceedance, 0), "x"), "")),
    color = "white", size = 3, fontface = "bold"
  ) +
  scale_fill_manual(values = risk_colors, name = "Risk Level") +
  labs(
    title = "Multi-Matrix Contamination Heatmap",
    subtitle = "Risk levels across sample types for top contaminated locations",
    x = "Sample Matrix",
    y = "Location",
    caption = "Numbers show maximum exceedance ratios > 10x"
  ) +
  theme_river_remedy() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

print(heatmap_plot)
```

# Conclusions and Recommendations

## Key Findings Summary

::: callout-important
## 🚨 Critical Environmental Crisis

The 2006 baseline assessment reveals a **catastrophic contamination scenario** in the Pilcomayo River Basin:

1.  **Systemic Contamination**: 90.4% of samples exceed WHO standards
2.  **Extreme Lead Pollution**: Up to 231,150x WHO limits in soil
3.  **Multi-Matrix Impact**: Contamination across water, soil, sediment, food, and biological systems
4.  **Geographic Hotspots**: Las Carreras, Huacareta, Camargo, and Yotala show extreme contamination
5.  **Human Health Crisis**: Elevated lead levels in human blood samples
:::

## Priority Action Areas

```{r priority-actions}
priority_actions <- tibble(
  Priority = c("Immediate", "Immediate", "Short-term", "Medium-term", "Long-term"),
  Action = c(
    "Emergency remediation at Huacareta and Las Carreras",
    "Public health intervention in affected communities", 
    "Source control and waste management improvements",
    "Comprehensive environmental restoration program",
    "Long-term monitoring and health surveillance"
  ),
  Timeline = c("0-3 months", "0-6 months", "6-12 months", "1-3 years", "Ongoing"),
  Estimated_Cost = c("High", "Medium", "High", "Very High", "Medium")
)

kable(priority_actions) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::row_spec(1:2, background = "#DC143C", color = "white")
```

## Regulatory Compliance

**Legal Framework Violations:** - Multiple violations of WHO drinking water guidelines - Exceedances of EPA soil contamination limits - Food safety concerns under Codex Alimentarius standards - Public health risks exceeding CDC blood lead reference values

## Future Research Priorities

1.  **Source Attribution**: Identify specific contamination sources
2.  **Temporal Trends**: Compare with 2011 Cerro Rico and 2024 TNC data
3.  **Bioaccumulation Studies**: Assess food chain contamination
4.  **Health Impact Assessment**: Comprehensive epidemiological study
5.  **Remediation Technology**: Evaluate treatment options

------------------------------------------------------------------------

*This report establishes the 2006 baseline for the River Remedy project's multi-temporal analysis of Pilcomayo Basin contamination. The findings underscore the urgent need for comprehensive environmental intervention and long-term monitoring.*

**Data Sources:** - Fundación I.T.A. (2006). Informe Final: Muestreo de aguas, suelos, vegetales, sangre humana y animales, peces y sedimentos en puntos seleccionados de la Cuenca del Río Pilcomayo - WHO Guidelines for Drinking-water Quality (2017) - Codex Alimentarius CXS 193-1995 Rev. 2018 - CDC Blood Lead Reference Value (2021)
